---
title: "2017fall semester_sleep analysis"
output:
  html_document: default
  pdf_document: default
---

For the past 85 days, I recorded when I went to bed and when I woke up. I then wanna use R to extract my own sleeping pattern from these data. This is just I play around with my data for fun. 

The way I recorded it causes many data problems and I need to fix them first. (For example, 8:00 recorded as 20:00; use Chinese ":" instead of English one ":")

```{r}
setwd("/Users/susan/Columbia/生活相关/Self-management")

if(!require(data.table)){install.packages(data.table)}
library(data.table)

sleep_data <- fread("2017fall学期睡眠分析 raw data.csv", na.strings = "")
colnames(sleep_data) <- c("week","date","bed","wake_up")
sleep_data <- sleep_data[complete.cases(sleep_data),]
```

###data cleaning 
```{r}
sleep_data$bed <- sub(pattern = "：", replacement = ":", x = sleep_data$bed)
sleep_data$wake_up <- sub(pattern = "：", replacement = ":", x = sleep_data$wake_up)

for (i in 6:11){
        y = i+12
        sleep_data$bed <- sub(pattern = paste(".*",i,":",sep=""), replacement = paste(y,":",sep=""), x = sleep_data$bed)
        
}
sleep_data$bed <- sub(pattern = paste(".*","12",":",sep=""), replacement = paste("0",":",sep=""), x = sleep_data$bed)

for (i in 16:23){
        y = i-12
        sleep_data$wake_up <- sub(pattern = paste(".*",i,":",sep=""), replacement = paste(y,":",sep=""), x = sleep_data$wake_up)
        
}
sleep_data$wake_up <- sub(pattern = paste(".*","24",":",sep=""), replacement = paste("12",":",sep=""), x = sleep_data$wake_up)
rm(y)
```

###data formating
```{r}
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)


sleep_data$date <- as.Date(sleep_data$date, format = "%m/%d")
sleep_data$bed <-  as.POSIXct(sleep_data$bed, format = "%H:%M") 
sleep_data$wake_up <-  as.POSIXct(sleep_data$wake_up, format = "%H:%M") 

```

##Sleep Analysis
```{r}
sleep_data$hour_of_sleep = 0

for (i in 1:nrow(sleep_data)){
if (unclass(sleep_data$bed[i]) >= 1515214800 & unclass(sleep_data$bed[i]) <= 1515240000) 
{sleep_data$hour_of_sleep[i] <- sleep_data$wake_up[i] - sleep_data$bed[i]} else 
{sleep_data$hour_of_sleep[i] <- sleep_data$wake_up[i] + 86400 - sleep_data$bed[i]}
}
rm(i)

sleep_data$hour_of_sleep <- round(sleep_data$hour_of_sleep, digits = 1)
```


**avergae hours of sleep every day**
```{r}
mean(sleep_data$hour_of_sleep)
```

**maximize hours of sleep in a day**
```{r}
max(sleep_data$hour_of_sleep)
```

**minimize hours of sleep in a day**
```{r}
min(sleep_data$hour_of_sleep)
```

variance of hours of sleep every day
```{r}
sd(sleep_data$hour_of_sleep)
```

##visualization 
```{r}
sleep_data$bed <- format(sleep_data$bed, format = "%H:%M") %>% as.factor()
sleep_data$wake_up <- format(sleep_data$wake_up, format = "%H:%M")  %>% as.factor()
```


time for bed (histogram by time)
```{r}
bed_time <- append(levels(sleep_data$bed)[17:26], levels(sleep_data$bed)[1:16])
sleep_data$bed <- factor(sleep_data$bed, levels = bed_time)
plot(table(sleep_data$bed),xlab = "time for bed (histogram by time)", ylab = "count")

```


time to wake up 
```{r}
plot(table(sleep_data$wake_up),xlab = "time to wake up (histogram by time)", ylab = "count")
```


hours of sleep 
```{r}
hist(sleep_data$hour_of_sleep[sleep_data$hour_of_sleep > 3],freq = TRUE, main = "sleeping hours frequency", ylim = range(0:8), xlab = "hour of sleep")
```

